### 2020-08-05 data/reports.tex
postdocs code, data; first generation

We have demonstrated the asymptotic analysis of a semi-linear optimal control problem posed on a smooth oscillating boundary domain in the present paper. We have considered a more general oscillating domain than the usual “pillar-type” domains. Consideration of such general domains will be useful in more realistic applications like circular domain with rugose boundary. We study the asymptotic behavior of the problem under consideration using a new generalized periodic unfolding operator. Further, we have studied the homogenization of a non-linear optimal control problem and such non-linear problems are limited in the literature despite the fact that they have enormous real-life applications. Among several other technical difficulties, the absence of a enough criteria for the optimal control is one of the most attention-grabbing issues in the current setting. We also obtain corrector results in this paper.  We considered an optimal control problem posed on a domain with a highly oscillating smooth boundary where the controls are applied on the oscillating part of the boundary. There are many results on domains with oscillating boundaries where the oscillations are pillar-type (non-smooth) while the literature on smooth oscillating boundary is very few. In this article, we use appropriate scaling on the controls acting on the oscillating boundary leading to different limit control problems; namely, boundary optimal control and interior optimal control problem. In the last part of the article, we visualize the domains as a branched structure, and we introduce unfolding operators to get contributions from each level at every branch.

@@

(1) A comprehensive, 40+ page paper treating distance matrices of trees. (2) An 80+ page survey on matrix positivity. (3) Extending results of Loewner, Horn, Cauchy, and Frobenius to new symmetric function identities and positivity conditions. (Among other works.)

@@

Multispecies versions of random juggling models were studied and exact stationary distributions were obtained using combinatorial techniques [1]. For one variant, ultrafast convergence to stationarity was proved.  A new totally asymmetric exclusion process with two species of particles on a finite one-dimensional lattice in contact with reservoirs was introduced, where the second-class particle is impermeable to the right reservoir [2]. The stationary distribution was obtained using a matrix ansatz and the nonequilibrium phase diagram was computed in the thermodynamic limit. This was later extended to multiple species of particles and the phase diagram was computed.  A new interacting particle system called the Hilbert-Galton board was introduced and various theorems about this process were proved [3].  Random walks on noncommutative finite rings generated by both addition and multiplication were studied. Results were proven about the spectrum, stationary distribution and mixing time [4].  A new class of factorizations about characters of the general linear group, when the variables were specialised, were proved [5]. Somewhat surprisingly, the resulting objects are themselves characters of other classical groups. Some applications to enumerations of plane partitions and alternating sign matrices was given.  The combinatorics of a large class of cyclic orders were studied [6]. In particular, the enumeration of circular extensions of these cyclic orders was related to the enumeration of a family of polytopes made popular by R. Stanley.  Extending earlier work, odd and chiral representations were studied for all Coxeter groups [7].   [1] A. Ayyer, C. Finn and D. Roy, Matrix product solution of a left-permeable two-species asym- metric exclusion process, Phys. Rev. E, 97 no. 1 (2018) 012151, 10pp. [2] A. Ayyer, J. Bouttier, S. Corteel, S. Linusson and F. Nunzi, Bumping sequences and multi- species juggling, Advances in Applied Mathematics, 98 (2018) 100–126. [3] A. Ayyer and S. Ramassamy, The Hilbert-Galton board, Latin American Journal of Probability and Mathematical Statistics (ALEA), 15 no. 2 (2018), 755–774. [4] A. Ayyer and P. Singla, Random motion on finite rings, II: Noncommutative rings,  arXiv:1807.04082. [5] A. Ayyer and R. E. Behrend, Factorization theorems for classical group characters, with applications to alternating sign matrices and plane partitions, arXiv:1804.04514 [6]  A. Ayyer, M. Josuat-Vergès and S. Ramassamy, Extensions of partial cyclic orders and consecutive coordinate polytopes,  arXiv:1803.10351. [7]  A. Ayyer, A. Prasad and S. Spallone, Macdonald trees and determinants of representations for finite Coxeter groups, arXiv:1812.00608.

@@

We have shown that the boundary of the pseudorhombicuboctahedron is the only semi-equivelar map on the 2-sphere which is not vertex-transitive.   As a consequence, we show that each semi-equivelar map on the real projective plane is vertex-transitive.

\subsection{Dilip P Patil}

(1) During my sabbatical leave in IIT Bombay I have written the following two manuscripts (jointly with Prof. J. K. Verma):  [1] (Jointly with Kriti Goel and J. K. Verma ) Nullstellens\"atze and Applications, to appear in Conference Proceedings Published by Springer.   In this expository paper, we present simple proofs of the Classical, Real, Projective  and Combinatorial Nullstellens\"atze. Several applications are also presented such as a classical theorem of Stickelberger for solutions of polynomial equations in terms of eigenvalues of commuting operators, construction of a principal ideal domain which is not Euclidean, Hilbert's $17^{th}$ problem,  the Borsuk-Ulam theorem in topology and solutions of the conjectures of Dyson, Erd\"{o}s and Heilbronn.  [2]  (Jointly wit  J. K. Verma) Rational points and generalized trace forms on a finite algebra over a real closed field, to appear in Conference Proceedings Published by Springer.   The core of this article is the equality of the number of $K$-rational points  with the signature of the trace form of a finite $K$-algebra over a real closed field $K$. The main tools are symmetric bilinear forms, hermitian forms, trace forms, generalized trace forms and their types and signatures.  Further, we prove a criterion for the existence of $K$-rational points by using generalized trace forms. As an application we prove the Pederson-Roy-Szpirglas theorem  about counting common real zeros of  real polynomial equations. We also prove deduce this theorem as a special case of a similar classical theorem for real closed fields.  [3] On the connectedness of the fiber product of two connected finte free affine schemes. Preprint.  This article grew out while studying Burnside algebra of finite abelian groups.

\subsection{E. K. Narayanan}

Study of Heckman-Opdam hypergeometric functions associated to root systems of type $BC,$ where the associated multiplicity function was allowed to assume negative values, was continued.

@@

Further work was undertaken to investigate a weak form of negative curvature --- known as visibility --- exhibited by a range of domains in $\mathbb{C}^n$ equipped with the Kobayashi distance. The class of domains that are visibility domains with respect to the Kobayashi distance was significantly broadened in the preprint [1]. Furthermore, two Wolff--Denjoy-type theorems were established for the latter class of domains.  [1] Gautam Bharali and Anwoy Maitra, A weak notion of visibility, a family of examples, and Wolff--Denjoy theorems.

@@

Detection of a causal relationship between two or more sets of data is an important problem across various scientific disciplines. Test statistics based on Granger causality ignore the effect of practical measurement impairments such as subsampling, additive noise, and finite sample effects. We modelled [1] the problem of detecting a causal relationship between two time series as a binary hypothesis test with the null and alternate hypotheses corresponding to the absence and presence of a causal relationship, respectively. We derived the distribution of the test statistic under the two hypotheses and show that measurement impairments can lead to suppression of a causal relationship between the signals, as well as false detection of a causal relationship, where there is none. We also used the derived results to propose two alternative test statistics for causality detection. These detectors are analytically tractable, which allowed us to design the detection threshold and determine the number of samples required to achieve a given missed detection and false alarm rate.   Prostate-associated gene 4 (PAGE4) is an intrinsically disordered protein implicated in prostate cancer. Experiments have shown that PAGE4 expands upon hyperphosphorylation and that this expansion is localized to its N-terminal half. To understand the interactions underlying this structural transition, we performed [2] molecular dynamics simulations using Atomistic AWSEM, a multi-scale molecular model that combines atomistic and coarse- grained simulation approaches. Our simulations show that electrostatic interactions drive transient formation of an N-terminal loop, the destabilization of which accounts for the dramatic change in size upon hyperphosphorylation. Finally, we construct a mechanism-based mathematical model that allows us to capture the interactions of different phosphoforms of PAGE4 with AP-1 and its downstream target, the androgen receptor (AR)—a key therapeutic target in prostate cancer. Our model predicts intracellular oscillatory dynamics indicating phenotypic heterogeneity in an isogenic cell population. Thus, conformational switching of PAGE4 may potentially affect the efficiency of therapeutically targeting AR activity.   We proposed [3] an application of the microtremor (ambient noise) H/V spectral ratio technique to identify significant rheological boundaries at shallow depths, estimate thickness of both lignite bearing Tertiary sedimentary sequence and late Cretaceous Deccan basalt flows and comprehend basinal geometry of Umarsar Basin (Babia syncline). Forty-six stations were gauged in a grid format at ~250 m resolution during the microtremor survey. We detected three rheological interfaces. Finally, we proposed the microtremor H/V spectral ratio technique as a tool to develop economical borehole plan with realistic reserve estimate and a step forward towards rapid economical assessment covering large mining lease areas.   [1] IEEE Trans Signal Proc. (2018) [2] J. Mol. Biol (2018) [3] J. Appl. Geophys. (2018)

@@

Working on certain problems on nodal sets of Gaussian random functions and in random matrix theory.

@@

We analyze an augmented political business cycle model taking into account the effect of employment creation decisions by the ruling party jointly on inflation and growth. The objective is to maximize voter support in the next election that depends on the rate of unemployment as well as that of growth and inflation. We allow for randomness in the New Keynesian Phillips Curve (NKPC) model for the relationship between  inflation and unemployment as well as in a benchmark labour productivity function for   analyzing the growth rate. We provide explicit solution paths of the affine Markov   control problem that results from our formulation. We also provide numerical illustrations with plausible parametric configurations to generate more insight into our model. [1] G. K. Basak, M. K. Ghosh and D. Mukherjee, A Stochastic Model with Inflation, Growth and Technology  for the Political Business Cycle, J. Computational Economics, 53 (2019), 125-140.

\subsection{Pooja Singla}

In [1], We study some irreversible Markov chains on finite commutative rings randomly generated using both addition and multiplication. We describe the eigenvalues and their multiplicities. We also prove recursive formulas for stationary distribution. We extend this work to finite non-commutative rings in [2].   In [3], the Gelfand-Whittaker model for $GL_n(O)$ and $SL_n(O)$,where $O$ is a  compact discrete valuation ring is studied. We proved that this model is multiplicity free and describe its constituents.   In [4] We describe the abscissa of convergence of representation zeta function of $SL_2(O)$ for a compact discrete valuation ring $O$ with even characteristic. For this case, we also give construction of continuous primitive irreducible representations and study group algebra isomorphism problem.   [1]. Random motion on finite rings, I: commutative rings, Arvind Ayyer and Pooja Singla, to appear in Algebras and Representation Theory.  [2]. Random motion on finite rings, II: Non-commutative rings; Arvind Ayyer and Pooja Singla, preprint.   [3]. A multiplicity one theorem for groups of type $A_n$ over discrete valuation rings, Shiv Prakash Patel and Pooja Singla, preprint.  [4]. Representation Growth of Special Compact Linear Groups of Order Two,  M. Hassain and Pooja Singla, preprint.

@@

By improving some earlier work of Felikson, Retakh and Tumarkin, a classification of all the regular subalgebras of Affind Kac-Moody Lie algebras have been obtained. By continuing this classification problem to Extend Affine Lie Algebras which are natural generalizaion of Affine Kac-Moody algebras, a classification of all the maximal closed sub-root systems of Affine reflection systems has also been done. The question of when two tensor products of irreducible integrable representations would be isomorphic to another for Borcherds-KacMoody algebras has been investigated.

@@

In [1], the critical behavior of the component sizes for the configuration model has been studied when the tail of the degree distribution of a uniform vertex is a regularly-varying function with exponent $\tau -1$, where $\tau\in (3, 4)$.   In [2], the metric space scaling limits of the minimal spanning trees of both the random (simple) 3-regular graph and the 3-regular configuration model have been established, and it has been shown that both these models belong to the universality class identified in the work of Addario-Berry, Broutin, Goldschmidt, and Miermont [Annals of Probability, 2017].   [1] Heavy-tailed configuration models at criticality.  Authors: Souvik Dhara, Remco van der Hofstad, Johan van Leeuwaarden, Sanchayan Sen.  Annales de l'Institut Henri Poincaré (B) Probabilités et Statistiques. Accepted in March 2019.  [2] Geometry of the minimal spanning tree of a random 3-regular graph.  Authors: Louigi Addario-Berry, Sanchayan Sen ArXiv preprint. Announced in October 2018.

@@

We refined our previous results where, as part of the PolyMath 14 project, we had classified homogeneous length functions on groups. In particular we showed that quasi-homogeneous length functions on groups with vanishing commutator length (such as solvable groups) are bounded distance from pullbacks from Abelianizations.  We also replicated and made robust the computer-assisited proof we had found in the course of the above mentioned PolyMath project and investigated general principles involved. We further developed methods for computer assisted theorem proving.

@@

We continue our investigations on analytic aspects of Automorphic forms.

\subsection{Srikanth K. Iyer}

The random connection model is studied in the connectivity regime. Strong law results for the critical scaling parameter required to eliminate isolated nodes is derived. A sufficient condition for the graph to be connected with high probability is obtained. Other results include strong law results for the maximum and minimum vertex degrees [1]. The capacity of a wireless cellular network is defined and an exact characterization (non-asymptotic) of this natural capacity metric is derived. It is shown that that the capacity increases polynomially with the base station (BS) density in the low BS density regime and then scales inverse exponentially with the increasing BS density [2]. A model of merge and split group dynamics, also called fission-fusion dynamics, for heterogeneous populations that contain two types (or species) of individuals is proposed and analysed. It is predicted using the analytical model that there is a critical group size below which groups are more likely to be homogeneous and contain the abundant type or species. Despite the propensity of heterogeneous groups to split at higher rates, we find that groups are more likely to be heterogeneous but only above the critical group size. Monte Carlo simulation of the model show excellent agreement with these analytical model results. The implications of our results to empirical studies on flocking systems are discussed [3].

@@

The main focus of my research during this period was in developing a correspondence between meromorphic quadratic differentials and certain geometric structures on punctured Riemann surfaces. Some of this work arose out of an effort to understand how such ``meromorphic" geometric structures arise in degenerations, in the context of hyperbolic structures. Other collaborative projects during this period that resulted in papers include a project with Basudeb Datta (IISc), on understanding geometric tilings of the hyperbolic plane, and work with Indranil Biswas (TIFR Mumbai) and Sorin Dumitrescu (CNRS, France) that concerns branched projective structures on Riemann surfaces.   Papers written or published during this period:  [1]   (with Michael Wolf)  Meromorphic quadratic differentials and measured foliations on a Riemann surface, Mathematische Annalen, Vol. 373 1-2,  (February 2019), pg. 73-118.  [2]  Limits of harmonic maps and crowned hyperbolic surfaces, accepted for publication , Transactions of the AMS,  available at  https://doi.org/10.1090/tran/7777.   [3]  Holomorphic quadratic differentials in Teichmüller theory,  to appear, Handbook of Teichmüller theory Vol VII, available at  http://arxiv.org/abs/1902.06406.  [4] (with Basudeb Datta)  Uniform tilings of the hyperbolic plane,  submitted, available at http://arxiv.org/abs/1806.11393.  [5] (with Indranil Biswas and Sorin Dumitrescu) Branched projective structures on a Riemann surface and logarithmic connections, submitted, available at http://arxiv.org/abs/1808.04555.

@@

We have obtained sparse bounds for the lacunary maximal function associated to spherical means on the Heisenberg group [1]. As a consequence we have obtained an analogue of a result of C. P. Calderon for the lacunary spherical maximal function on Euclidean spaces.

>>>>

The main theme of the research is to understand the asymptotic analysis of  PDE problems and related optimal control problems in some complex domains. We consider domains whose boundary is rapidly oscillating which  leads to multi-scales in the problem. One of the major issue is the development of the so called unfolding operators which contains information about both scales. We have ingeniously used the unfolding operators not only to carry out the limiting behaviour of the problem under study, it is used to characterize the optimal controls. One needs to construct suitable such operators according to the type of domain and the problem [1], [3-5]. On a completely different direction, a multi-step Volterra integral equation-based algorithm was developed to measure the electric field auto-correlation function from multi-exposure speckle contrast data. This enabled us to derive an estimate of the full diffuse correlation spectroscopy data-type from a low-cost, camera-based system. This method is equally applicable for both single and multiple scattering field auto-correlation models. The feasibility of the system and method was verified using simulation studies, tissue mimicking phantoms and subsequently in in vivo experiments [2].

@@

(1) I extended a fundamental matrix inequality, the Schur Product Theorem, for positive definite matrices, by providing nonzero, tight lower bounds. (2) I improved and updated several preprints - notably, a comprehensive work titled "Moment-sequence transforms" with Belton-Guillot-Putinar.

@@

In joint work with Aas, Linusson and Potka [1], the exact phase diagram for a semipermeable TASEP with nonlocal boundary jumps was obtained. In joint work with Behrend [2], new factorization theorems for classical group characters were proved. Applications to alternating sign matrices and plane partitions were given. In joint work with Finn and Roy [3], the phase diagram for a multispecies left-permeable asymmetric exclusion process was derived. In joint work with Linusson [4], a new class of reverse juggling processes was defined and their stationary distributions were computed.

@@

A semi-regular tiling of the hyperbolic plane is a tessellation by regular geodesic polygons with the property that each vertex has the same vertex-type. Determined combinatorial criteria for the existence, and uniqueness, of a semi-regular tiling with a given vertex-type \ref{bdsg_2019} .

\subsection{Dilip P. Patil}

COMPUTATIONAL ASPECTS OF BURNSIDE RINGS PART II: IMPORTANT MAPS (Jointly Martin  Kreuzer)   The Burnside ring B(G) of a finite group G, a classical tool in the representation theory of finite groups, is studied from the point of view of computational algebra. In the rst part (cf. [8]) we examined the ring theoretic properties of B(G) using the methods of computer algebra. In this part we shift our focus to important maps between two Burnside rings and make several well-known maps of representation theory explicitly computable. The inputs of all our algorithms are the tables of marks of the two groups, and the outputs are matrices of integers representing the maps via their image in the ghost ring. All algorithms have been implemented in the computer algebra system ApCoCoA and are illustrated by applying them to explicit examples. Especially, we study the restriction and induction maps, the projection and in ation maps, the conjugation isomorphism, and the Frobenius-Wielandt homomorphism.

\subsection{E K Narayanan}

Positivity properties and asymptotics of Heckman-Opdam hypergeometric functions associated to root system to type BC with non-positive multiplicity functions were studied in detail.

@@

A detailed investigation of the Bergman geometry of domains whose boundaries have isolated infinite-type points was completed. Optimal estimates, from above and below, for the growth of the Bergman kernel (on the diagonal) and the Bergman metric, as one approaches an infinite-type boundary point, were obtained. The class of domains that were studied subsumes all classes of domains that have previously been considered in the literature. These results have appeared in the paper [1].  [1] Bharali, Gautam, On the growth of the Bergman metric near a point of infinite type, J. Geom. Anal. , 30 (2020), 1238-1258.

@@

We constructed [1] symmetry preserving and symmetry broken N-bright, dark and antidark soliton solutions of a nonlocal nonlinear Schrödinger equation. To obtain these solutions, we used appropriate eigenfunctions in Darboux transformation (DT) method. Further, we derived two dark/antidark soliton solution with the help of DT method. In the dark/antidark soliton solution case we observed a contrasting behaviour between the envelope of the ﬁeld and parity transformed complex conjugate envelope of the ﬁeld. For a particular parametric choice, we got dark (antidark) soliton for the ﬁeld while the parity transformed complex conjugate ﬁeld exhibits antidark (dark) soliton. Due to this surprising result, both the ﬁeld and PT transformed complex conjugate ﬁeld exhibit sixteen different combinations of collision scenario.   We presented observational evidence [2] from studies on primary cortical cultures from AD transgenic mice for significant decrease in total spine density at DIV-15 and onward. This indicates reduction in potential healthy synapses and strength of connections among neurons. Based on this, a network model of neurons was developed, that explains the consequent loss of coordinated activity and transmission efficiency among neurons that manifests over time. The critical time when structural connectivity in the brain undergoes a phase-transition, from initial robustness to irreparable breakdown, is estimated from this model. We also showed how the global efficiency of signal transmission in the network decreases over time. Moreover, the number of multiple paths of high efficiency decreased rapidly as the disease progresses, indicating loss of structural plasticity and inefficiency in choosing alternate paths or desired paths for any pattern of activity. Thus loss of spines caused by β-Amyloid (Aβ) peptide results in disintegration of the neuronal network over time with consequent cognitive dysfunctions in Alzheimer’s Disease (AD).  We investigated [3] the effect of four-wave mixing in the interactions among nonlinear waves such as solitons, breathers, and rogue waves of a coupled generalized nonlinear Schrodinger equation. We explored several interesting results including superposition of breather pulses, increment in the number of breather pulses and in amplitudes of breathers, and rogue waves. By strengthening the four-wave mixing parameter, we observed different transformations that occur between different localized structures. For instance, we visualized a transformation from bright soliton to breather form, bright and dark rogue wave to four-petaled rogue wave structures, four-petaled rogue wave to other rogue wave forms, and so on. Another important observation was that the interaction of a bright soliton with a rogue wave in the presence of the four-wave mixing effect provides interaction between a dark oscillatory soliton and a rogue wave.  [1] Vishnu Priya, N., Senthilvelan, M.,  Rangarajan, G. & Lakshmanan, M. On symmetry preserving and symmetry broken bright, dark and antidark soliton solutions of nonlocal nonlinear Schrödinger equation, Physics Letters A, 383, 15 (2019).   [2] Kashyap, G. A. R. S. R. K., Bapat, D., Das, D., Gowaikar, R. D., Amritkar, R. E., Rangarajan, G., Ravindranath, V. & Ambika, G., Synapse loss and progress of Alzheimer’s disease - A network model, Scientific Reports, 9, 6555 (2019).   [3] Vishnu Priya, N., Senthilvelan, M. &  Rangarajan, G, On the role of four-wave mixing effect in the interactions between nonlinear modes of coupled generalized nonlinear Schrodinger equation, Chaos, 29, 123135 (2019).

\subsection{Hrish Seshadri}

\subsection{Kaushal Verma}

For a given pair of smoothly bounded planar domains $D_1, D_2$, it was shown in \cite{1} that the product of the Carath\'{e}odory metrics on them is bounded from below by the product of the Carath\'{e}odory metrics on their union and intersection up to a multiplicative factor. This shows that the S\"{z}ego kernels on $D_1, D_2, D_1 \cap D_2$ and $D_1 \cup D_2$ are related by the same inequality.    The Suita conjecture in higher dimensions is a paradigm that asks for the limiting boundary values of the product of the Bergman kernel (on the diagonal) and the volume of the Kobayashi indicatrix. This object is a biholomorphic invariant and is known to admit universal bounds on convex domains. Its limiting behaviour was clarified on strongly peudoconvex domains and egg domains in $\mathbb C^2$. See \cite{2}.  The boundary behaviour of several classical conformal invariants were studied in \cite{3}.  The relationship between the Fridman invariant and the Squeezing function was clarified in \cite{4}.  The relation between a pair of H\'{e}non maps that share the same forward and backward Julia sets was clarified in \cite{5}.

\subsection{Mahesh Kakde}

Jointly with Samit Dasgupta from Duke University I resolved the Brumer-Stark conjecture during the said period. We spoke about it at the "International Colloquium" in January 2020 in TIFR. Further, a paper on higher Chern classes written a while ago (jointly with Frauke Bleher, Ted Chinburg,  Ralph Greenberg, George Pappas, Romyar Sharifi and Martin Taylor) was accepted for publication in American Journal of Mathematics.

@@

Completed a study of law of iterated logarithm for last passage percolating on the two dimensional integer lattice [1]. Other studies on absolute continuity of limiting spectral distributions, a proof of KMT theorems, study of discrete nodal length, a problem on amplitude constrained Gaussian channel and some problems on nodal domains of random functions are in various stages of completion.  [1] Lower Deviations in β-ensembles and Law of Iterated Logarithm in Last Passage Percolation.   Authors: Riddhipratim Basu, Shirshendu Ganguly, Milind Hegde, Manjunath Krishnapur. arXiv:1909.01333

@@

Risk-sensitive stochastic differential games have been studied in the nonnegative orthant [1] and the whole space [2]. Existence of saddle point/Nash equilibria for relevant cases have been derived.

\subsection{Purvi Gupta}

Note: this description is for research carried out during the period February 11, 2020 (date of joining) to March 31, 2020.   In this period,  I completed and submitted a manuscript ([1]),  based on joint work with C. U. Wawrzyniak. This work studies the stability of the solution of the so-called global Bishop problem (a plateau-type problem) for a model real $n$-dimensional sphere in $\mathbb{C}^n$, $n\geq 2$. In this problem, the solution is required, among other things, to be foliated by holomorphic disks (i.e., Levi-flat) attached to the given sphere. This structure is significant in complex analysis for multiple reasons. The general Bishop problem, although very well-understood in two (complex) dimensions, is wide open in dimensions three and higher, and our work is the first global result in this direction.    References [1] Gupta, Purvi and Wawrzyniak, Chloe Urbanski, Stability of the hull(s) of an $n$-sphere in $\mathbb{C}^n$. Submitted.

\subsection{R VENKATESH}

(a) Settled the decomposition theorem for G-stable affine Demazure modules for the cases G = E_6,7,8 and F_4 and thus completed the main theorems of the papers of V. Chari and et al., (J. Algebra, 2016) and Deniz Kus and et al., (Represent. Theory., 2016). Obtained a new combinatorial proof for the key fact that was used in  V. Chari and et al., (J. Algebra, 2016) to prove the decomposition of G-stable affine Demazure modules. This key fact was proved only for the cases when G is the classical type and G_2 in their paper and they also proved it by analysing case-by-case. Our proof is uniform avoids the case-by-case analysis and works for all finite-dimensional simple Lie algebras.  (b) Simplified the presentation of all affine Demazure modules of untwisted affine Kac–Moody Lie algebras. Simplifying the generators and relations of these modules is very crucial in their study. We are investigating how to use this simplified presentation of affine Demazure modules to study Schur positivity type questions which arise naturally in Algebraic Combinatorics.   (c) A number of conjectures made about the graded characters affine Demazure modules of affine sl_3 using data obtained from SAGE.

@@

(a) Settled the decomposition theorem for G-stable affine Demazure modules for the cases G = E_6,7,8 and F_4 and thus completed the main theorem of the paper of V. Chari and et al., (J. Algebra, 2016). Obtained a new combinatorial proof for the key fact that was used in  V. Chari and et al., (J. Algebra, 2016) to prove the decomposition of G-stable affine Demazure modules. This key fact was proved by V. Chari and et al., (J. Algebra, 2016) only for the cases when G is the classical types and G_2 and they did a case-by-case analysis to prove this. Our new proof is very uniform and avoids the case-by-case analysis and works in general.  (b) Simplified the presentation of all affine Demazure modules of untwisted affine Kac–Moody Lie algebras. Simplifying the generators and relations of these modules is very crucial in their study. We are investigating how to use this simplified presentation of affine Demazure modules to study Schur positivity type questions which arise naturally in Algebraic Combinatorics.   (c) A number of conjectures made about the graded characters affine Demazure modules of affine sl_3 using data obtained from SAGE.

\subsection{RADHIKA GANAPATHY}

1) Let $G$ be a split connected reductive group over $\mathbb{Z}$. Let $F$ be a non-archimedean local field. With $K_m := Ker(G(O_F ) → G(O_F /p_{m,F} ))$, Kazhdan proved that for a field $F′$ sufficiently close to $F$, the Hecke algebras $\mathcal{H}(G(F), K_m)$ and $\mathcal{H}(G(F′), K_m′ )$ are isomorphic, where $K_m′$ denotes the corresponding object over $F′$. I have generalized this result to general connected reductive groups [1].  2) In my earlier work, I had established  a variant of the Kazhdan isomorphism for split reductive groups (by replacing $K_m$ replaced by the $m$-th filtration subgroup of the Iwahori subgroup). I have been working on generalising this result to general connected reductive groups. As part of this work, I have chosen a nice set of representatives in $G(F)$ for the elements of the extended affine Weyl group. This set of representatives has some very interesting properties (for instance, it is compatible with unramified descent), and is useful in the study of representation theory of $p$-adic groups.   [1] A Hecke algebra isomorphism over close local fields, preprint, 12 pages.

@@

In [1], we provide an explicit algorithm for sampling a uniform simple connected random graph with a given degree sequence. By products of this central result includes scaling limits for the metric space structure of the maximal components in the critical regime of both the configuration model and the uniform simple random graph model with prescribed degree sequence under finite third moment assumption on the degree sequence. As a substantive application we answer a question raised by Černý and Teixeira, Random Structures Algorithms (2013), by obtaining the metric space scaling limit of maximal components in the vacant set left by random walks on random regular graphs.  In [2], we study limits of the largest connected components (viewed as metric spaces) obtained by critical percolation on uniformly chosen graphs and configuration models with heavy-tailed degrees. We develop general principles under which the scaling limits can be obtained. Of independent interest, we derive refined asymptotics for various susceptibility functions and the maximal diameter in the barely subcritical regime.  In [3], we give alternate constructions of (i) the scaling limit of the uniform connected graphs with given fixed surplus, and (ii) the continuum random unicellular map of a given genus that start with a suitably tilted Brownian continuum random tree and make `horizontal' point identifications, at random heights, using the local time measures. In particular, this yields a breadth-first construction of the scaling limit of the critical Erdős-Rényi random graph which answers a question posed by Addario-Berry, Broutin, and Goldschmidt, Probab. Theory Related Fields (2012). As a consequence of this breadth-first construction we obtain descriptions of the radii, the distance profiles, and the two point functions of these spaces in terms of functionals of tilted Brownian excursions.    [1] Geometry of the vacant set left by random walk on random graphs, Wright's constants, and critical random graphs with prescribed degrees Authors: Shankar Bhamidi, Sanchayan Sen Journal: Random Structures & Algorithms (First published on 25 July 2019)  [2] Universality for critical heavy-tailed network models: Metric structure of maximal components Authors: Shankar Bhamidi, Souvik Dhara, Remco van der Hofstad, Sanchayan Sen Journal: Electronic Journal of Probability (Accepted in December 2019)  [3] On breadth-first constructions of scaling limits of random graphs and random unicellular maps Authors: Grégory Miermont, Sanchayan Sen ArXiv preprint https://arxiv.org/abs/1908.04403 (Announced in August 2019)

@@

We continued our development of automated theorem proving systems. Specifically, we developed a framework based on equation terms to allow accumulation of knowledge and facilitate deep learning and hybrid autonomous/interactive execution.

@@

Work on the determination of Siegel modular forms by their fundamental Fourier coefficients was done including a mod p version. Study of Fourier coefficients and sup-norms of modular forms was continued.

\subsection{Srikanth Krishnan Iyer}

We study combinatorial connectivity for two models of random geometric complexes. These two models - \v{C}ech and Vietoris-Rips complexes - are built on a homogeneous Poisson point process of intensity $n$ on a $d$-dimensional torus, $d > 1$, using balls of radius $r_n$. Given a (simplicial) complex (i.e., a collection of $k$-simplices for all $k \geq 1$), we can connect $k$-simplices via $(k+1)$-simplices (`up-connectivity') or via $(k-1)$-simplices (`down-connectivity). Our interest is to understand these two combinatorial notions of connectivity for the random \v{C}ech and Vietoris-Rips complexes asymptotically as $n \to \infty$. Our results give tighter bounds on the constants in the logarithmic scale as well as shed light on the possible second-order correction factors. Further, they also reveal interesting differences between the phase transition in the \v{C}ech and Vietoris-Rips cases. The analysis is interesting due to non-monotonicity of the number of isolated $k$-faces (as a function of the radius) and leads one to consider `monotonic' vanishing of isolated $k$-faces. The latter coincides with the vanishing threshold mentioned above at a coarse scale (i.e., $\log n$ scale) but differs in the $\log \log n$ scale for the \v{C}ech complex with $k = 1$ in the up-connected case. For the case of up-connectivity in the Vietoris-Rips complex and for $r_n$ in the critical window, we also show a Poisson convergence for the number of isolated $k$-faces when $k \leq d$. [1]

@@

(1) With Mahan Mj (TIFR)  we completed a  research program concerning meromorphic projective structures on surfaces and their monodromy representations. We used new  work of Allegretti-Bridgeland, together with older ideas of Thurston, to develop a geometric understanding of these structures on punctured surfaces, and prove analogues of results of Hejhal and Gallo-Kapovich-Marden -- see [1] and [2]. In an independent paper [3], I used our the techniques we had developed, to shed light on a remark of Poincaré in 1883. This complemented other projects concerning meromorphic quadratic differentials (see [8] and [10]).  (2) In joint work [4] with Indranil Biswas (TIFR), Mahan Mj (TIFR) and Junho Whang (MIT), we classified the finite mapping class group orbits in the PSL(2,C) character variety.  (3) In a project with Basudeb Datta (IISc) we revised and published a paper [5] developing an algorithmic technique of constructing semi-regular tilings of the hyperbolic plane. We expect this to have an impact on tilings-related research, which is an active area that has an intersection with several fields like combinatorics, computational geometry and group theory.   (4)  With Harish Seshadri (IISc) we wrote a paper [6], and a survey [7], concerning the complex geometry of Teichmüller spaces, where we investigated the convexity (or lack thereof) of the boundary of complex domains biholomorphic to Teichmüller space. We have ongoing conversations regarding other questions that lie at the intersection of several complex variables, differential geometry and Teichmüller theory.   (5) A paper with Indranil Biswas (TIFR) and Sorin Dumitrescu (UCA)  concerning branched projective structures on surfaces was published [9]. In an ongoing project, we aim to understand the symplectic geometry of the space of such structures better.  [1] Meromorphic projective structures, grafting and the monodromy map, submitted, available at http://arxiv.org/abs/1904.03804. [2] Monodromy representations of meromorphic projective structures, Proc. Amer. Math. Soc. 148 (2020), no. 5, 2069–2078. [3] Monodromy groups of CP1-structures on punctured surfaces, submitted, available at https: //arxiv.org/abs/1909.10771. [4] Surface group representations in SL(2,C) with finite mapping class orbits, submitted, available at https://arxiv.org/ abs/1707.00071. [5] Semi-regular tilings of the hyperbolic plane, to appear, Discrete and Computational Geometry (2019), available at https://doi.org/10.1007/s00454-019-00156-0. [6] On domains biholomorphic to Teichmüller spaces, International Mathematics Research Notices, Volume 2020, Issue 8, (2020), 2542–2560. [7] Complex geometry of Teichmüller domains, Handbook of Teichmüller theory Vol VII, 63–88, IRMA Lect. Math. Theor. Phys., 30, Eur. Math. Soc., Zürich, 2020. [8] Holomorphic quadratic differentials in Teichmüller theory, Handbook of Teichmü ller theory Vol VII, 2020. [9] Branched projective structures on a Riemann surface and logarithmic connections, Documenta Mathematica 24 (2019), 2299–2337. [10] Limits of harmonic maps and crowned hyperbolic surfaces, Trans. Amer. Math. Soc. 372 (2019), no. 11, 7573–7596.

@@

In a joint work with Pritam Ganguly  we have investigated the $ L^p $  boundedness of the lacunary maximal function $ M_{\H^n}^\lac $ associated to the spherical means $ A_r f$  taken over Koranyi spheres on the Heisenberg group. Closely following an approach used by M. Lacey in the Euclidean case, we obtain sparse bounds for these maximal functions leading to new unweighted and weighted estimates. The key ingredients in the proof are the $L^p$  improving property of the operator $ A_r $  and a continuity property of the difference $ A_r f−\tau_y A_r f,$ where $\tau_yf(x)=f(xy−1) $ is the right translation operator.

\subsection{Thirupathi Gudi}

We have developed a Hybrid High Order methods for elliptic obstacle problem in both two and three space dimensions allowing general meshes. The gap in literature of developing an optimal order convergent method using quadratic finite element method is resolved under the realistic regularity. In another work, we derived optimal order finite element analysis for constrained Dirichlet boundary optimal control problem. We have also studied numerous finite element methods for parabolic obstacle problem under the realistic regularity.
